Molecular dynamics simulation of synchronization in driven particles

### 2020-09-27 manuscript/manuscript_ajp_2020.tex
changed last sentence of abstract
outlined details, started introduction

  %What problem did you study and why is it important?
  We study synchronization of interacting particles
  driven by an externally applied force while confined to a narrow channel. 
  %What methods did you use?
  With
  numerical simulations, we control the particle interactions,
  external force and particle environment
  in order to mimick experimental studies of driven colloidal particles
  confined by light-fields.
  We use molecular dynamics simulations to model the particle dynamics,
  using overdamped equations of motion suitable for a viscous suspension
  of microscopic particles.
  %What were your principal results?
  We study 
  particle synchronization under a variety of conditions,
  including static and dynamic patterns formed on
  periodic and quasiperiod substrates,
  kink propagation on these surfaces,
  and the limits of applied force that
  cause the particles to transition to chaotic behavior.
  %What conclusions have you drawn from these results?
  We demonstrate the transition from
  trapped to sliding dynamics in
  quasiperiodic landscapes differs
  from that of a periodic landscape by [TBD],
  and that kinks of high density propagate
  [more/less ?  ] when [what?].
  We include sample code and opportunties
  for students to reproduce our results and propose
  new numerical experiments.

@@

\section{Introduction} % Section titles are automatically converted to all-caps.
% Section numbering is automatic.

Synchronization is one of the most diverse fundamental physical phenomena. Synchronization has been studied for over three centuries, from Huygens pendulum clocks to the rhythmic beats of the flapping wings in a flock of birds. It is a goal of physicists to discover synchronization in different forms and learn how to control the phenomenon that occurs when weakly coupled competing oscillators adjust rhythms to match each other. A synchronization phenomenon that the physics community is particularly interested in is dynamic phase locking, which occurs when naturally oscillating motions are driven by an external modulation \cite{juniper2015}. Another interest of ours is to study the transition from ordered, synchronized dynamics, to disordered, chaotic dynamics. Chaotic dynamics is the irregular, unpredictable time evolution of nonlinear systems and occurs in mechanical oscillators \cite{chaos}. Here, we analyze the microscopic dynamics of phase locking in a colloidal model system using molecular dynamics (MD).

\section{Molecular Dynamics Model}

The simulations use overdamped dynamics with Euler's method in a classical two-dimensional model for particles confined to a periodic potential. With only a few particles in two-dimensions, the simulation runs quickly, making this an appropriate model for undergraduates to explore.

@@

\section{MD Simulations}

@@

\subsection{Multi-particle system}

We also simulated a twenty particle system confined to a narrow channel, as shown in Fig. 2a). The interparticle forces of neighboring particles cause the system to form a buckled chain when the system is annealed. When a single particle is driven, the neighboring particles act similarly to a periodic landscape to impede its motion. A driven particle can exhibit mode locking with a well-chosen AC drive and frequency. In the attached movie, Figure2.mp4, we show the complex dynamics of mode locking, where the driven particle leap-frogs past the other particles.

@@

\section{Quasiperiodic substrate}
\label{DispEqSection}	% You can label sections for reference

>>>>

  %What problem did you study and why is it important?
  We study synchronization of interacting particles
  confined to a narrow channel 
  driven by an externally applied force.  
  %What methods did you use?
  With
  numerical simulations, we control the particle interactions,
  external force and particle environment
  in order to mimick experimental studies of driven colloidal particles
  confined by light-fields.
  We use molecular dynamics simulations to model the particle dynamics,
  using overdamped equations of motion suitable for a viscous suspension
  of microscopic particles.
  %What were your principal results?
  We study 
  particle synchronization under a variety of conditions,
  including static and dynamic patterns formed on
  periodic and quasiperiod substrates,
  kink propagation on these surfaces,
  and the limits of applied force that
  cause the particles to transition to chaotic behavior.
  %What conclusions have you drawn from these results?
  We demonstrate the transition from
  trapped to sliding dynamics in
  quasiperiodic landscapes differs
  from that of a periodic landscape by [TBD],
  and that kinks of high density propagate
  [more/less ?  ] when [what?].
  We include sample code
  and exercises for students
  that include 
  opportunties
  to reproduce our results and propose
  new numerical experiments.

@@

\section{Introduction} %
%

%Outline
%describe types of physical systems 
Numerical simulations of confined, driven particles
can describe a diverse variety of physical phenomena.
Particles which interact over long distances
include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more detail and references]
Particles which interact over short distances include
bubble arrays/emulsions [more systems and references].

Narrow channels studies are useful to provide insights 
of how particles move through geometries
such as wires, microchannels, and biological systems such as neuron axons and capillaries [more detail and references].
Stabilized charged particles form patterns
due to the interplay of the confining environment
and particle interactions.
Many such systems execute local oscillations
about stable points [elaborate].
%
An applied external force
increases the diversity of behaviors,
and can cause particles to flow in
a variety of non-linear complex behaviors
including
synchronized, aperiod, or chaotic dynamical patterns.
%
The presence of a modulating surface
can modify these patterns in a variety of ways,
changing the onset of dynamical flows,
and the overall flow patterns.

Colloidal particles trapped in light fields have proven
a particularly useful medium for studying these behaviors.%cite 
The relatively large size of the colloids
and ease in control has lead to a rich array of 
experimental results. %THIS IS A TERRIBLE SENTENCE.
Such studies are considered model systems
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE

Synchronization has been studied for over three centuries,
and is observed in different forms
from Huygens pendulum clocks
to the rhythmic beats of the flapping wings in a flock of birds.
Controlling
synchronization phenomena in weakly coupled oscillators
can be achieved with an external driving force
that causes the syncing of natural oscillation frequencies,
dynamic phase locking \cite{juniper2015}.
[useful and interesting because?]

Disordered chaotic dynamics are also possible,
where irregular, unpredictable time evolution of
nonlinear systems and occurs in mechanical oscillators \cite{chaos}.

In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD},
including various particle interactions and confining environments.
In Section~\ref{sec:sync} we
demonstrate how uniform environments and applied forces
and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams.
We modify the environment and applied forces
and show aperiodic, or nearly periodic flow behaviors
in Section~\ref{sec:quasiperiod}.
Finally we explore the transition to chaos in 
in Section~\ref{sec:chaos}.

%Here we model the microscopic dynamics of phase locking in a colloidal model system using molecular dynamics (MD).

\section{Molecular Dynamics Model}
\label{sec:MD}
The simulations use overdamped dynamics with Euler's method in a classical two-dimensional model for particles confined to a periodic potential. With only a few particles in two-dimensions, the simulation runs quickly, making this an appropriate model for undergraduates to explore.

@@

%\section{MD Simulations}

@@

\subsection{Synchronization in multi-particle systems}
\label{sec:sync}

We simulated a twenty particle system confined to a narrow channel, as shown in Fig. 2a). The interparticle forces of neighboring particles cause the system to form a buckled chain when the system is annealed. When a single particle is driven, the neighboring particles act similarly to a periodic landscape to impede its motion. A driven particle can exhibit mode locking with a well-chosen AC drive and frequency. In the attached movie, Figure2.mp4, we show the complex dynamics of mode locking, where the driven particle leap-frogs past the other particles.

@@

\section{Quasiperiodic substrate}
\label{sec:quasiperiod}	% You can label sections for reference

\section{Chaotic dynamics}
\label{sec:chaos}	% You can label sections for reference
